\documentclass[draft]{agujournal2019}
\usepackage{amsmath,amssymb}
\usepackage{booktabs,longtable,array,tabularx}
\usepackage{url,xurl}
\usepackage{hyperref}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setlength{\emergencystretch}{3em}
\sloppy
% Pandoc's longtable wrapper asks for a caption type even when the Markdown
% table has no caption; AGU leaves that type as ``none``.  Define only the
% otherwise-missing counter so the official class can typeset the table.
\makeatletter
\@ifundefined{c@none}{\newcounter{none}}{}
\providecommand{\thenone}{\arabic{none}}
\providecommand{\fnum@none}{Table~\thenone}
\makeatother
\DeclareUnicodeCharacter{00B0}{\ensuremath{^\circ}}
\DeclareUnicodeCharacter{00B1}{\ensuremath{\pm}}
\DeclareUnicodeCharacter{00B2}{\textsuperscript{2}}
\DeclareUnicodeCharacter{00B3}{\textsuperscript{3}}
\DeclareUnicodeCharacter{00B7}{\ensuremath{\cdot}}
\DeclareUnicodeCharacter{00D7}{\ensuremath{\times}}
\DeclareUnicodeCharacter{2013}{--}
\DeclareUnicodeCharacter{2014}{---}
\DeclareUnicodeCharacter{207B}{\textsuperscript{-}}
\DeclareUnicodeCharacter{2212}{\ensuremath{-}}
\DeclareUnicodeCharacter{2264}{\ensuremath{\leq}}
\DeclareUnicodeCharacter{2265}{\ensuremath{\geq}}

\journalname{Water Resources Research}

\begin{document}

\title{ThermoRoute: a bounded-deviation, physics-inspired framework for retrospective daily river-temperature hindcasting}

\authors{[Verified authors and affiliations required before submission]}
\affiliation{1}{[Department / Laboratory, Institution, City, Country]}
\correspondingauthor{[Verified corresponding author]}{[verified.email@example.org]}

\begin{keypoints}
\item A frozen protocol separates development, model freeze, predictor evidence, and one-time outcome opening
\item Primary comparisons use identical station/date/horizon keys and whole-HUC2 clustered inference
\item This pre-opening draft reports methods and limitations; current empirical results are intentionally absent
\end{keypoints}

\begin{abstract}
Daily river water temperature is difficult to hindcast honestly because persistence
is strong, issue-time information must be separated from future observations, sensor
records are incomplete, and nearby sites are not statistically independent. We
describe ThermoRoute, a physics-inspired statistical predictor that combines a
damped-persistence anchor, a learned thermal-relaxation proposal, a sparse
horizon-conditioned variable/lag router, a causal temporal convolutional encoder,
a regime mixture, a bounded point residual, separate three-quantile heads, and a
split-conformal 90\% interval. The
canonical development data contain 657,480 site-days from 120 stable USGS site
numbers during 2006--2020, spanning 34 states and 15 HUC2 groups. Years 2019--2020
are development/exploratory data. A repository-internal pre-label protocol defines
a later one-time retrospective interval, six primary model classes, five formal
station-balanced comparisons, exact whole-HUC2 sign-flip p-values, whole-HUC2
cluster-bootstrap intervals, and Holm multiplicity control. The protocol also
freezes model/input chronology, raw outcome-quality evidence, and deterministic
claim rendering. The current repository has not yet completed the canonical model
rerun, model-suite freeze, predictor-evidence freeze, authorization, or opening.
Accordingly, no empirical performance conclusion is made here. The contribution at
this stage is a testable architecture and a fail-closed evaluation system whose
remaining scientific and engineering limitations are stated explicitly.

\textbf{Keywords:} river water temperature; retrospective hindcasting; physics-inspired
machine learning; clustered inference; conformal prediction; reproducibility.
\end{abstract}

\section{Problem and scope}\label{problem-and-scope}

Water-temperature predictions may be useful for scientific monitoring and planning,
but evaluation is easy to inflate. Daily temperature often changes slowly, so a
copy of the last observation can be difficult to beat. Weather and flow variables
can also leak future information if their timestamp or issue-time availability is
not defined. Finally, rows from the same station and region are dependent; treating
all daily rows as independent creates implausibly small uncertainty estimates.

This study asks a narrow empirical question: on fixed common station/\allowbreak date/\allowbreak horizon
keys, can a constrained learned correction improve on strong statistical references
and remain competitive with LightGBM and a global LSTM? The study separately asks
whether empirical intervals and event probabilities are useful descriptions of
uncertainty. It does not infer an energy budget, hydraulic state, or management
utility from the architecture.

The repository contains an older three-station cascade case study and a larger USGS
development panel. Those legacy outputs are not used as current evidence because
they were produced before stable USGS site numbers, the 120v2 panel contract, and
prediction-lineage sidecars were enforced.

\subsection{Scientific context and related work}\label{scientific-context-and-related-work}

River thermal regimes reflect atmospheric forcing, discharge, groundwater,
shading, channel form, and regulation; a statistical predictor does not identify
that heat budget (\href{https://doi.org/10.1111/j.1365-2427.2006.01597.x}{Caissie, 2006}).
Earlier approaches range from nonlinear air--water regressions
(\href{https://doi.org/10.1029/98WR01877}{Mohseni et al., 1998}) to hybrid air-temperature
and discharge models (\href{https://doi.org/10.1088/1748-9326/10/11/114011}{Toffolon and Piccolroaz, 2015})
and broader statistical comparisons (\href{https://doi.org/10.1002/hyp.10913}{Piccolroaz et al., 2016}).
Recent work evaluates machine learning, deep sequence models, spatial transfer,
and physics-guided river-network models
(\href{https://doi.org/10.5194/hess-25-2951-2021}{Feigl et al., 2021};
\href{https://doi.org/10.1088/1748-9326/abd501}{Rahmani et al., 2021};
\href{https://doi.org/10.1137/1.9781611976700.69}{Jia et al., 2021};
\href{https://doi.org/10.3389/frwa.2023.1184992}{Zwart et al., 2023}). ThermoRoute\textquotesingle s
narrow contribution is a common-key, clustered, pre-opening benchmark of a
bounded correction; it is not a claim to replace these process or network models.

The implementation uses USGS NWIS daily values, Daymet meteorology, and gridMET
wind (\href{https://doi.org/10.5066/F7P55KJN}{USGS NWIS};
\href{https://doi.org/10.3334/ORNLDAAC/2129}{Daymet V4};
\href{https://doi.org/10.1002/joc.3413}{Abatzoglou, 2013}). LightGBM is the principal
tree baseline (\href{https://proceedings.neurips.cc/paper/2017/hash/6449f44a102fde848669bdd9eb6b76fa-Abstract.html}{Ke et al., 2017}),
and interval calibration follows conformalized quantile regression
(\href{https://papers.nips.cc/paper/2019/hash/5103c3584b063c431bd1268e9b5e76fb-Abstract.html}{Romano et al., 2019}).

\section{Data}\label{data}

\subsection{Canonical development panel}\label{canonical-development-panel}

The canonical file is \texttt{data\_\allowbreak{}usgs/\allowbreak{}panel\_\allowbreak{}usgs\_\allowbreak{}120v2.parquet}, bound to
\texttt{data\_\allowbreak{}usgs/\allowbreak{}station\_\allowbreak{}registry\_\allowbreak{}v1.csv} by \texttt{data\_\allowbreak{}usgs/\allowbreak{}frozen\_\allowbreak{}panel\_\allowbreak{}v1.json}.
Each of 120 sites has one row for every calendar day from 2006-01-01 through
2020-12-31, giving 657,480 rows. The registry contains 34 states and 15 HUC2
groups. The raw columns consumed by the model, in frozen order, are water
temperature (\texttt{WTEMP}), streamflow (\texttt{FLOW}), air temperature (\texttt{TEMP}), precipitation
(\texttt{PRCP}), a relative-humidity proxy (\texttt{RHMEAN}), Daymet daylight-period mean
incoming shortwave radiation in W m⁻² (\texttt{DH}), and wind speed (\texttt{WDSP}). \texttt{DH} is not
day length, and \texttt{RHMEAN} is derived from temperature and vapor pressure rather than
a direct daily-mean humidity observation. Water level may be retained as raw
evidence but is not a model input.

Observed-value missingness in the frozen panel is approximately 15.8\% for \texttt{WTEMP},
2.8\% for \texttt{FLOW}, about 0.07\% for \texttt{TEMP}, \texttt{PRCP}, \texttt{RHMEAN}, and \texttt{DH}, and zero for
\texttt{WDSP}. These rates describe the committed derived panel; they do not reconstruct
provider request history. The small Daymet gap is structured rather than random:
all 120 sites lack the provider-calendar final day in leap years 2008, 2012, 2016,
and 2020. \texttt{WLEVEL}, retained only as raw evidence, is missing on 77.65\% of rows.

The pre-opening environmental audit records 1,345 rejected candidate stations in
the legacy selection ledger (950 without joint WTEMP/FLOW availability, 376 with
low full-period coverage, and 19 with low 2019--2020 development-period coverage).
Among retained stations, 38 share a repeated HUC code with another retained
station and 19 have another retained station within 10 km. WTEMP and FLOW have
maximum missing runs of 2,404 and 1,369 days, respectively. Two stations contain
2,059 signed negative-FLOW rows (minimum −121 cfs), whose tidal, backwater, or
measurement semantics are not resolved. These facts reinforce that the cohort is
availability-enriched and spatially dependent, rather than nationally
representative.

The surviving station ledgers are consistent with full-period WTEMP/FLOW coverage
thresholds of 0.55/0.70 and 2019--2020 thresholds of 0.80/0.80. Consequently,
2019--2020 participated in cohort construction as well as later model and narrative
development. The original discovery responses, retrieval timestamps, exact command
line, and complete run configuration were not retained. The original 1,465-candidate
selection is therefore auditable from committed ledgers but is not source-replayable.
The development panel also retains no NWIS qualifier columns, method or sensor
history, or original daily-value response bytes; measurement discontinuities and
qualifier sensitivity cannot be reconstructed for 2006--2020.

The temporal roles are fixed before fitting:

{\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.12}
\noindent
\begin{tabularx}{\dimexpr\linewidth-24pt\relax}{@{}lrrr>{\raggedright\arraybackslash}X@{}}
\toprule\noalign{}
Role & Dates & Rows & Observed \texttt{WTEMP} & Permitted use \\
\midrule\noalign{}
Training & 2006--2015 & 438,240 & 341,646 & fit preprocessing and models \\
Validation & 2016--2017 & 87,720 & 84,074 & choose frozen settings \\
Calibration & 2018 & 43,800 & 42,279 & CQR and event-reference calibration \\
Development evaluation & 2019--2020 & 87,720 & 85,621 & previously inspected; exploratory diagnosis only \\
\bottomrule\noalign{}
\end{tabularx}
}

The 2019--2020 partition was inspected and informed cohort, model, and narrative
development; it is exploratory and not independently confirmatory.

\subsection{Target-period inputs and outcomes}\label{target-period-inputs-and-outcomes}

The target interval is fixed in the protocol. Historical Daymet and gridMET
covariates must be retrieved and archived before outcome access. Meteorology is
represented at each station coordinate, not aggregated over upstream catchments.
The primary information set uses values dated on or before each historical issue
date in retrospectively retrieved, finalized products and consumes no
horizon-specific future weather forecast. This is a date-indexed retrospective
hindcast; it does not establish that the same predictor vintages were available
operationally at that time.

Because the legacy panel lacks its original HTTP responses, a separate pre-opening
bridge must re-fetch 2018--2020 Daymet/gridMET with the confirmation parser and
compare every predictor on the exact site/date registry. This gate can detect
product-version, parsing, scaling, missingness, and one-day alignment drift. It
cannot prove that provider calendar days share NWIS local-day boundaries, nor can
it reconstruct as-issued data availability; those remain permanent limitations.

After the final local protocol seal, we added this outcome-free 2018--2020
parser/product compatibility gate. The post-seal engineering gate strengthens
reproducibility but was not part of the sealed protocol and does not alter its
hypotheses, estimands, model registry, margins, or decision rules. Formal training
remains blocked until the offline comparison publishes
\texttt{PASS\_\allowbreak{}EXACT\_\allowbreak{}PRODUCT\_\allowbreak{}BRIDGE}.

The raw-only opening child is restricted to USGS daily-value requests for daily
mean water temperature, streamflow, and water level. It records exact request and
response bytes, series identifiers, qualifiers, final URLs, timestamps, and
hashes. If simultaneous finite series disagree, the normalized value is missing
and every constituent is retained in the quality audit. No site can be replaced
after outcomes are accessed.

\subsection{Metadata-only external cohort}\label{metadata-only-external-cohort}

After the model-suite commit, a fixed 34-state metadata query identifies stream
sites advertising daily water-temperature capability without requesting values or
coverage dates. A deterministic seed selects 30 stations whose exact USGS site IDs
do not occur in the development registry. This is site-ID disjointness, not proof
of HUC8, river-network, upstream/downstream, or distance separation. Models for
this cohort are pooled and station-agnostic where applicable. The exercise remains
outside the five-test family and cannot change model or site selection.

\section{Model design}\label{model-design}

\subsection{Conservative anchor and learned proposal}\label{conservative-anchor-and-learned-proposal}

Let \texttt{y\_\allowbreak{}t} be the last observed daily water temperature and \texttt{c\_\{t+h\}} a frozen
seasonal climatology. Damped persistence moves \texttt{y\_\allowbreak{}t} toward \texttt{c\_\{t+h\}} at a fixed
rate and provides the principal anchor. ThermoRoute also produces a flow- and
season-conditioned relaxation proposal. That proposal is learned from data; its
parameter is not interpreted as a measured residence time or heat-transfer
coefficient.

\subsection{Variable/lag router, TCN, and regime mixture}\label{variablelag-router-tcn-and-regime-mixture}

For each horizon, a sparse router assigns weights over the seven variables at
lags 0--14. The sequence builder supplies a 32-day tensor, but this is a
construction buffer rather than an effective-memory claim: the current
two-block, kernel-three causal temporal convolutional network has a theoretical
seven-step receptive field, while the router's oldest usable value is lag 14.
Thus no input older than lag 14 can affect the current model output. A
mixture-of-experts combines the resulting representations. The router is an
allocation mechanism inside the predictor, not a map of connected reaches.
These components follow sparsemax attention
(\href{https://proceedings.mlr.press/v48/martins16.html}{Martins and Astudillo, 2016}),
generic temporal convolutional sequence modeling
(\href{https://arxiv.org/abs/1803.01271}{Bai et al., 2018}), and sparsely gated expert
mixtures (\href{https://arxiv.org/abs/1701.06538}{Shazeer et al., 2017}).

\subsection{Bounded residual and probabilistic heads}\label{bounded-residual-and-probabilistic-heads}

The learned point correction is transformed by \texttt{tanh} and scaled by a fixed
\texttt{delta\_\allowbreak{}scale}, limiting its numerical deviation from the anchor. This yields a
finite algebraic bound relative to that reference for the point output. It does
not bound absolute error, event-tail error, interval width, or behavior after a
distribution shift.

The learned models emit a separate MSE point forecast and pinball-trained 0.05,
0.50, and 0.95 quantile heads. Split conformalized quantile regression fitted on
2018 widens only the nominal q05--q95 interval; q50 is not CQR-adjusted. Ensemble
quantiles are member-wise averages, not quantiles of a mixture distribution.
Temporal CQR offsets are station-by-horizon, whereas external offsets are pooled
by horizon. A separate event head uses each temporal station\textquotesingle s 2006--2015 q90, or
one pooled development-training q90 in the external arm, and one Platt calibrator
per horizon fitted on 2018. Probability summaries give every retained station
equal total weight and claim empirical marginal coverage only. Quantile crossing,
station-balanced achieved coverage, interval width, three-quantile pinball mean,
Brier score, log loss, discrimination, reliability, and calibration slope/intercept
are descriptive outputs.

\section{Evaluation design}\label{evaluation-design}

\subsection{Primary models and common keys}\label{primary-models-and-common-keys}

The six primary model types are persistence, damped persistence, climatology,
LightGBM, a global LSTM, and ThermoRoute. The three deterministic references have
no seed; LightGBM, LSTM, and ThermoRoute use seeds 0--4. LightGBM selects among four
candidate settings separately by horizon on 2016--2017; the LSTM selects among
three candidate architectures with seed 0 before fitting five members. These
tuning budgets are documented but not identical. The
temporal mechanism audit additionally freezes seven one-factor controls:
prior-only, no dynamic prior, fixed relaxation, no router, no mixture, no TCN, and
unbounded residual. Every comparison uses identical station/\allowbreak date/\allowbreak horizon target
keys for the two models.

The temporal arm is future-time hindcasting at known gauges. ThermoRoute and the
LSTM receive a stable station embedding, and temporal LightGBM receives the same
site identity as a categorical feature. The external pooled arm disables all
station-identity inputs and uses only pooled development-training transformations,
but still consumes each target site\textquotesingle s observed WTEMP history through the issue
date. It is therefore a history-dependent new-gage evaluation and cannot be
interpreted as prediction without target-site water-temperature records.

An admissible issue must have genuinely observed issue-date WTEMP. Other cells in
the 32-day history may be filled with training-only seasonal medians and retain
explicit missingness masks; no minimum observed-history fraction is required.
Development windows require all three target horizons to remain within one
partition and have observed labels, whereas later confirmation evaluates horizon
admissibility independently. FLOW uses signed \texttt{log1p}, preserving negative values
rather than clipping them or calling them drought; PRCP uses nonnegative \texttt{log1p}.
Temporal scaling and imputation are station-specific and train-only, while the
external arm uses pooled train-only transformations.

The current canonical configuration sets \texttt{delta\_\allowbreak{}scale} to 1.0 °C. It is a
development-selected algebraic point bound relative to the damped anchor, not an
error, tail-risk, or safety bound. Earlier 2019--2020 inspection informed project
development, so this setting is not presented as selection from an untouched
development test.

For each station and horizon, RMSE is calculated only when at least 100 paired
targets are valid for both models. The primary effect is the unweighted median of
station-level candidate-minus-reference RMSE. Daily rows therefore do not directly
determine the between-station weight.

\subsection{Formal five-test family}\label{formal-five-test-family}

The frozen family contains:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  ThermoRoute versus damped persistence at 1 day, margin 0.00 °C.
\item
  ThermoRoute versus damped persistence at 3 days, margin 0.00 °C.
\item
  ThermoRoute versus damped persistence at 7 days, margin 0.00 °C.
\item
  ThermoRoute versus LightGBM at 3 days, numerical ceiling +0.05 °C.
\item
  ThermoRoute versus LightGBM at 7 days, numerical ceiling +0.05 °C.
\end{enumerate}

Each raw one-sided p-value is obtained by enumerating every whole-HUC2 sign
configuration for the frozen reportable cohort. A 10,000-draw whole-HUC2 cluster
bootstrap provides the interval for the median station effect. Holm adjustment is
applied to exactly these five p-values
(\href{https://www.jstor.org/stable/4615733}{Holm, 1979}). A favorable formal statement is allowed
only when the row is estimable, the Holm p-value is at most 0.05, and the interval
upper bound is strictly below the frozen margin. Disagreement between the p-value
and interval rules is reported conservatively. The 1-day LightGBM comparison is
descriptive only.

The sign-flip procedure assumes independent whole-cluster sign reversals and
joint sign symmetry of each complete HUC2 effect vector around the tested margin.
Exact enumeration removes Monte Carlo error but does not make this assumption
distribution-free. Before reportability attrition, the 15 HUC2 groups contain
2--26 stations, the largest contains 21.7\% of stations, and the inverse-Herfindahl
effective cluster count is 9.54. The percentile cluster-bootstrap interval is
therefore interpreted cautiously. Effective-cluster, largest-share, and
leave-one-HUC diagnostics carry \texttt{NO\_\allowbreak{}STRONG\_\allowbreak{}INFERENCE} when their predeclared
small/unequal-cluster warnings fire; they do not change the sealed five-test rule.

\subsection{Secondary analyses}\label{secondary-analyses}

Predeclared secondary outputs include exact-approved-qualifier target sensitivity,
equal-HUC and leave-one-HUC influence summaries, station-balanced probability
diagnostics, architecture controls, and synthetic missingness/noise/flow/weather
perturbations. They cannot promote or replace a failed formal test. A hypothetical
cost--loss calculation, if retained, is an illustration rather than observed
decision impact.

The optional Air2stream-style a4/a8 implementation is an exploratory physical
reference variant fitted by deterministic six-start bounded least squares on
measured drivers. It is not the official Air2stream implementation, and no result
against it supports a claim against the published model. A published-model
comparison requires the official implementation and documented calibration search.

\section{Reproducibility and one-time opening}\label{reproducibility-and-one-time-opening}

The final protocol bytes are bound to both their original Git preregistration
commit and the final pre-label commit by SHA-256. The evidence explicitly lacks an
external timestamp or independent custodian. The claim ledger derives phase only
from the canonical authorization, intent marker, and a fully verified receipt.

Before authorization, a fresh isolated process must replay all trained members,
heads, transforms, development predictions, dependency versions, runtime probe,
and source hashes without reading target-period outcomes. Git chronology must show
that the model-suite commit predates candidate metadata and target-period predictor
artifacts. Any source change after the model freeze invalidates authorization.

The opening creates one irreversible intent and one fixed request ledger. A network
continuation may fetch only ledger entries whose transaction directory is wholly
absent, under the same opening identifier. Partial or inconsistent transactions are
not replaced. After normalization or trusted publication begins, automatic
continuation is prohibited by the sealed protocol. The trusted scorer writes
create-only predictions, audits, all five tests, a report, and a self-hashed receipt.

\section{Current progress}\label{current-progress}

Completed repository work includes the canonical panel/registry seal, the final
pre-label protocol, deterministic model serialization contracts, development replay
code, exact clustered inference, probability/event contracts, outcome-quality
provenance, a dual-profile release verifier, and a structured claim ledger.

Required work still outstanding at this manuscript state is:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  complete the offline 2018--2020 predictor bridge and publish
  \texttt{PASS\_\allowbreak{}EXACT\_\allowbreak{}PRODUCT\_\allowbreak{}BRIDGE};
\item
  rerun the canonical six-model development experiment, the seven mandatory
  single-seed controls, and the separate matched plain-neural/feature-ladder audit;
\item
  freeze and independently replay every five-member temporal and pooled external
  learned-model bundle;
\item
  commit the model-suite registry before obtaining external metadata and target-period
  predictor artifacts;
\item
  freeze those outcome-free inputs and validate the Git chronology receipt;
\item
  run the full preflight suite, including the implemented same-opening transport
  continuation, before creating the unique authorization;
\item
  execute the fixed opening, render all five receipt-derived statements,
  regenerate artifacts, and verify the clean-room release.
\end{enumerate}

Until those steps finish, all legacy numeric outputs are considered stale and no
current empirical performance conclusion is available.

\section{Permanent limitations}\label{permanent-limitations}

The following machine-verifiable statements are part of the evidence contract and
remain in the manuscript regardless of the eventual numerical outcome.

Route A is history-dependent and uses target-site water-temperature observations through each issue date; it does not establish ungauged prediction.

Route A does not establish independent river-network separation, physical river-network routing, or hydraulic travel time.

Route A is a one-shot retrospective historical-information evaluation, not an operational replay with archived as-issued predictor vintages or future NWP.

Route A uses daily-mean statistical thresholds and a numerical non-inferiority margin; neither has ecological, biological, or regulatory meaning.

Route A provides no physical, deployment, regulatory, or distribution-free safety guarantee, and failure to reject is not equivalence.

Route-A architecture controls and predictor sensitivities are descriptive or exploratory and do not identify causal mechanisms.

Route A evaluates a fixed availability-enriched 120-site cohort and is not nationally representative of all U.S. rivers or all calendar days.

Route A does not establish conditional coverage, and its equal-weight three-quantile pinball summary is not CRPS.

Additional limitations are the availability-enriched station sample; missing
original provider bytes for the development panel; point-scale meteorology rather
than upstream forcings; unbalanced and coarse HUC2 clusters; daily-mean outcomes;
outcome-observability conditioning; a numerical margin without stakeholder-derived
importance; retrospective finalized covariates; absence of latency, memory, energy,
and multi-hardware benchmarks; owner-controlled local evidence chronology; the
unreplayable original 1,465-candidate discovery execution; and the absence of an
official Air2stream calibration run.

\section{Results placeholder and stopping rule}\label{results-placeholder-and-stopping-rule}

No formal result block is present in this pre-opening manuscript. After a valid
receipt exists, \texttt{scripts/\allowbreak{}26\_\allowbreak{}validate\_\allowbreak{}claims.py} will derive five deterministic
blocks from the exact statistics artifact. Each must appear exactly once, including
negative, conflicting, or non-estimable outcomes. Handwritten substitutions are
rejected by body hashes and evidence bindings.

\section{Data and code availability}\label{data-and-code-availability}

The canonical derived panel, stable station registry, HUC metadata snapshot, code,
protocol, and tests are in this repository. Original HTTP bytes and retrieval
timestamps for the 2006--2020 panel are unavailable, so replay begins from the
committed Parquet artifact. Candidate, meteorological, and target-outcome requests
are designed to retain exact raw evidence in a completed release. A DOI, public
repository URL, data license, and verified author metadata have not yet been
assigned and must be completed before submission.

\section{Conclusion}\label{conclusion}

ThermoRoute is presently a specified research system, not a completed empirical
finding. The repository has fail-closed controls for leakage,
cohort identity, model replay, clustered inference, raw provenance, statement
generation, and release closure. Its scientific value will depend on a complete
canonical rerun and the one-time evidence chain. The implementation is not ready
for confirmatory opening until every listed data, bridge, replay, chronology,
authorization, and release gate passes; the old headline numbers do not represent
the current repository.

\end{document}
